R Notebook

### 2021-07-11 draft implementation 2.Rmd
Delete some drafts
modifica

---
header-includes: 
- \newcommand{\R}{\textnormal{\sffamily\bfseries R}}
- \def\code#1{\texttt{#1}}
- \newtheorem{algorithm}{Algorithm}[section]
title: "R Notebook"

@@

\
\textbf{Implementation}
Consider again the random walk plus noise model of section XX,the challenge here is to estimate the filtered states using a Sequential Importance Sampling algorithm. The main idea of the SIS applied to this univariate linear gaussian model is described in the following algorithm. We indicate with $n$ the sample size of time observations and with $N$ the generated sample size for each step of the Sequenial Monte Carlo. 
\begin{algorithm} SIS filter for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\end{enumerate}
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{itemize}
\end{algorithm}
Our \code{SIS} function implemented in \R replicate this steps.
\
\hrule
\hrule
\code{SIS(data,N,m0,C0,tau,sigma)}
\hrule
\textbf{Arguments}
\
\code{data} \ \ the observed process. It has to be a vector or a univariate time series.\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{tau} \ \ the standard deviation $\tau$ in Equation (2)
\
\code{sigma} \ \ the standard deviation $\sigma$ in Equation (1)
\hrule
\hrule
```{r}
SISfun<-function(data,N,m0,C0,tau,sigma){
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)
  for(t in 1:length(data)){
    x    = rnorm(N,x,tau)                   #sample from N(x_{t-1},tau)
    w    = w*dnorm(data[t],x,sigma)         #update weight
    xs = rbind(xs,x)
    ws = rbind(ws,w)

@@

\textbf{Implementation}
In this section, Bootstrap Particle Filter will be used to estimate filtered states of the Random Walk plus Noise introduced in section XX. The overall strategy replicate the SIS filter with the addition of a ESS-based resampling step that applies when the effective sample size is smaller than a predetermined threshold opportunely chosen (in our example we decide to follow a common rule of thumb consisting in setting the threshold at $N/2$). The steps are presented in the following algorithm.
\begin{algorithm} BPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\end{enumerate}
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{itemize}
\end{algorithm}
These steps are resumed in our \code{PF} function.
\
\hrule
\hrule
\code{PF(data,N,m0,C0,tau,sigma,r)}
\hrule
\textbf{Arguments}
\
\code{data} \ \ the observed process. It has to be a vector or a univariate time series.\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{tau} \ \ the standard deviation $\tau$ in Equation (2)
\
\code{sigma} \ \ the standard deviation $\sigma$ in Equation (1)
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
```{r}
PFfun<-function(data,N,m0,C0,tau,sigma,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

@@

ggarrange(p1,c1,p2,c2,p3,c3,nrow=3,ncol=2,labels=c("N=100","","N=1000","","N=10000",""))
```
\
\textbf{Implementation}
\
The Bootstrap Particle Filter Approach for the Random Walk plus Noise Model described in Section XX can be improved accounting for the observations in the importance transition density and it consists of generating $x_{t}$ from its conditional distribution given $x_{t-1}$ and $y_{t}$. In the Normal model we are considering, the optimal proposal will be a Normal density as well with mean and variance given by 
\begin{align*}
\mu_{opt}=E(x_{t}|x_{t-1},y_{t})&=x_{t-1}+\frac{\tau^{2}}{\tau^{2}+\sigma^{2}}(y_{t}-x_{t-1})\\
\sigma_{opt}^{2}=V(x_{t}|x_{t-1},y_{t})&=\frac{\tau^{2}\sigma^{2}}{\tau^{2}+\sigma^{2}}
\end{align*}
On the other hand, the incremental weights, using this importance transition density, are proportional to the conditional density of $y_{t}$ given $x_{t-1}=x_{t-1}^{(i)}$ , i.e $N(x_{t-1}^{(i)},\tau^{2}+\sigma^{2})$, evaluated at $y_{t}$. In other words the algorithm implemented in \R is:
\begin{algorithm} GPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item Compute $\sigma_{opt}^{2}$ 
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $\mu_{opt}$
\item Draw $x_{t}^{(i)} \sim N(\mu_{opt},\sigma_{opt}^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2+\tau^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\end{enumerate}
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{itemize}
\end{algorithm}
The \code{GPF} function resume this passages.
\
\hrule
\hrule
\code{GPF(data,N,m0,C0,tau,sigma,r)}
\hrule
\textbf{Arguments}
\
\code{data} \ \ the observed process. It has to be a vector or a univariate time series.\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{tau} \ \ the standard deviation $\tau$ in Equation (2)
\
\code{sigma} \ \ the standard deviation $\sigma$ in Equation (1)
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
```{r}
GPFfun<-function(data,N,m0,C0,tau,sigma,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  importancesd<-sqrt(tau - tau^2 /(tau + sigma))
  predsd <- sqrt(sigma+tau)
  w  = rep(1/N,N)

@@

    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)
  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```
Let's provide directly a brief comparison between the Bootstrap Particle Filter (BPF) and the Guided Particle Filter (GPF). As we can see from the Figure XX, the Guided Particle Filter provides point estimates that are slightly better with respect to the ones of the BPF, and this is confirmed by the Table showing the RMSE. Moreover, also the variance is better, suggesting for higher precision.

@@

pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
gpf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)

@@

>>>>

---
header-includes: 
- \newcommand{\R}{\textnormal{\sffamily\bfseries R}}
- \def\code#1{\texttt{#1}}
- \newtheorem{algorithm}{Algorithm}[section]

title: "R Notebook"

@@

\
\textbf{Implementation}
Consider again the random walk plus noise model of section XX,the challenge here is to estimate the filtered states using a Sequential Importance Sampling algorithm. The main idea of the SIS applied to this univariate linear gaussian model is described in the following algorithm. We indicate with $n$ the sample size of time observations and with $N$ the generated sample size for each step of the Sequenial Monte Carlo. 
\begin{algorithm} SIS filter for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
Our \code{SIS} function implemented in \R replicate this steps.
\
\hrule
\hrule
\code{SIS(data,N,m0,C0,tau,sigma)}
\hrule
\textbf{Arguments}
\
\code{data} \ \ the observed process. It has to be a vector or a univariate time series.\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{tau} \ \ the standard deviation $\tau$ in Equation (2)
\
\code{sigma} \ \ the standard deviation $\sigma$ in Equation (1)
\hrule
\hrule
```{r}
SISfun<-function(data,N,m0,C0,tau,sigma){
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)
  for(t in 1:length(data)){
    x    = rnorm(N,x,tau)                   #sample from N(x_{t-1},tau)
    w    = w*dnorm(data[t],x,sigma)         #update weight
    xs = rbind(xs,x)
    ws = rbind(ws,w)

@@

\textbf{Implementation}
In this section, Bootstrap Particle Filter will be used to estimate filtered states of the Random Walk plus Noise introduced in section XX. The overall strategy replicate the SIS filter with the addition of a ESS-based resampling step that applies when the effective sample size is smaller than a predetermined threshold opportunely chosen (in our example we decide to follow a common rule of thumb consisting in setting the threshold at $N/2$). The steps are presented in the following algorithm.
\begin{algorithm} BPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ form $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item Draw $x_{t}^{(i)} \sim N(x_{t-1}^{(i)},\tau^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
These steps are resumed in our \code{PF} function.
\
\hrule
\hrule
\code{PF(data,N,m0,C0,tau,sigma,r)}
\hrule
\textbf{Arguments}
\
\code{data} \ \ the observed process. It has to be a vector or a univariate time series.\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{tau} \ \ the standard deviation $\tau$ in Equation (2)
\
\code{sigma} \ \ the standard deviation $\sigma$ in Equation (1)
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
```{r}
PFfun<-function(data,N,m0,C0,tau,sigma,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

@@

ggarrange(p1,c1,p2,c2,p3,c3,nrow=3,ncol=2,labels=c("N=100","","N=1000","","N=10000",""))
```
\
\textbf{Implementation}
\
The Bootstrap Particle Filter Approach for the Random Walk plus Noise Model described in Section XX can be improved accounting for the observations in the importance transition density and it consists of generating $x_{t}$ from its conditional distribution given $x_{t-1}$ and $y_{t}$. In the Normal model we are considering, the optimal proposal will be a Normal density as well with mean and variance given by 
\begin{align*}
\mu_{opt}=E(x_{t}|x_{t-1},y_{t})&=x_{t-1}+\frac{\tau^{2}}{\tau^{2}+\sigma^{2}}(y_{t}-x_{t-1})\\
\sigma_{opt}^{2}=V(x_{t}|x_{t-1},y_{t})&=\frac{\tau^{2}\sigma^{2}}{\tau^{2}+\sigma^{2}}
\end{align*}
On the other hand, the incremental weights, using this importance transition density, are proportional to the conditional density of $y_{t}$ given $x_{t-1}=x_{t-1}^{(i)}$ , i.e $N(x_{t-1}^{(i)},\tau^{2}+\sigma^{2})$, evaluated at $y_{t}$. In other words the algorithm implemented in \R is:
\begin{algorithm} GPF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item Compute $\sigma_{opt}^{2}$ 
\item For $t=1,...,n$:
\begin{enumerate}
\item Compute $\mu_{opt}$
\item Draw $x_{t}^{(i)} \sim N(\mu_{opt},\sigma_{opt}^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t-1})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t-1})$
\item Set $w_{t}^{(i)} = w_{t-1}^{(i)}f_{N}(y_{t};x_{t}^{(i)},\sigma^2+\tau^{2}) \ \ i=1,...,N$ such that $\{(x_{t},w_{t})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{t}|y_{t})$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
The \code{GPF} function resume this passages.
\
\hrule
\hrule
\code{GPF(data,N,m0,C0,tau,sigma,r)}
\hrule
\textbf{Arguments}
\
\code{data} \ \ the observed process. It has to be a vector or a univariate time series.\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{tau} \ \ the standard deviation $\tau$ in Equation (2)
\
\code{sigma} \ \ the standard deviation $\sigma$ in Equation (1)
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule
```{r}
GPFfun<-function(data,N,m0,C0,tau,sigma,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  importancesd<-sqrt(tau - tau^2 /(tau + sigma))
  predsd <- sqrt(sigma+tau)
  w  = rep(1/N,N)

@@

    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)
  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```

```{r echo=FALSE}
gpf<-GPFfun(y,N,m0,C0,tau,sigma)
ESS.GPF<-ESSplot(df,gpf)
```

```{r echo=FALSE, fig.align='center', fig.cap="a) GPF Filtered States with credible interval (in red). b) Effective sample size (in black) with threshold (in yellow).", fig.height=4, fig.pos='ht', fig.width=8, message=FALSE, warning=FALSE, results=F}
FILT.GPF<-Filtplot(df,gpf,"")
ggarrange(FILT.GPF,ESS.GPF,labels=c("(a)","(b)"))
```

Let's provide directly a brief comparison between the Bootstrap Particle Filter (BPF) and the Guided Particle Filter (GPF). As we can see from the Figure XX, the Guided Particle Filter provides point estimates that are slightly better with respect to the ones of the BPF, and this is confirmed by the Table showing the RMSE. Moreover, also the variance is better, suggesting for higher precision.

@@

pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
gpf2<-GPFfun(y,N,m0,C0,tau,sigma,r=k)

@@

For illustration purposes, we are going implement an auxiliary particle filter with auxiliary function $g(x_{t-1})=E(x_{t}|x_{t-1})=x_{t-1}$.  
\begin{algorithm} APF for Random Walk plus Noise Model
\begin{itemize}
\item Let $\{(x_{0},w_{0})^{(i)}\}_{i=1}^{N}$ summarizes $p(x_{0}|y_{0})$ such that, for example, $E(g(x_{0})|y_{0}) \approx \sum_{i=1}^{N}w_{0}^{(i)}g(x_{0}^{(i)})$. In particular, initialize $(x_{0}^{(1)},...,x_{0}^{(N)})$ from $N(m_{0},C_{0})$ and set $w_{0}^{(i)}=N^{-1} \ \forall \ i=1,...,N$.
\item For $t=1,...,n$:
\begin{enumerate}
\item For $k=1,...,N$:
\begin{enumerate}
\item Draw $I_{k}$ with $P(I_{k}) \propto w_{t-1}^{(i)}f(y_{t}|g(x_{t-1}^{(i)})) $
\item Draw $x_{t}^{(k)} \sim N(x_{t-1}^{(I_{k})},\tau^2)$
\item Set  $\tilde{w}_{t}^{(k)} = \frac{f_{N}(y_{t}|x_{t}^{(k)})}{f_{N}(y_{t}|g(x_{t-1}^{(I_{k})}))}$
\end{enumerate}
\item Normalize the weights: $w_{t}^{(i)}=\frac{\tilde{w}_{t}^{(i)}}{\sum_{j=1}^{N}(\tilde{w}_{t}^{(j)})}$
\item Compute $ESS=\Bigg(\sum_{i=1}^{N}(w_{t}^{(i)})^{2}\Bigg)^{-1}$
\item if $ESS<N/2$ then
\begin{enumerate}
\item Draw a sample of size N, $(x_{t}^{(1)},...,x_{t}^{(N)})$, from the discrete distribution $P(x_{t}=x_{t}^{(i)})=w_{t}^{(i)},\ \ i=1,...,N$
\item Reset the weights: $w_{t}^{(i)}=N^{-1}$, $i=1,...,N$.
\end{enumerate}
\item Set $p(x_{t}|y_{t})=\sum_{i=1}^{N}w_{t}^{(i)}\delta_{x_{t}^{(i)}}$
\end{enumerate}
\end{itemize}
\end{algorithm}
The \code{APF} function resume this passages.
\
\hrule
\hrule
\code{APF(data,N,m0,C0,tau,sigma,r)}
\hrule
\textbf{Arguments}
\
\code{data} \ \ the observed process. It has to be a vector or a univariate time series.\
\code{N} \ \ number of particles generated at each step
\
\code{m0} \ \ central value of the normal prior state distribution
\
\code{C0} \ \ variance of the normal prior state distribution
\
\code{tau} \ \ the standard deviation $\tau$ in Equation (2)
\
\code{sigma} \ \ the standard deviation $\sigma$ in Equation (1)
\
\code{r} \ \  if present the threshold is set equal to $N/r$ otherwise, if missing, the threshold is set equal to $N/2$
\hrule
\hrule

```{r}
APFfun<-function(data,N,m0,C0,tau,sigma,r){
  if(missing(r)){r=2}else{}
  xs<-NULL
  ws<-NULL
  ess<-NULL
  x  = rnorm(N,m0,sqrt(C0))
  w  = rep(1/N,N)

    weight = w*dnorm(data[t],x,sigma)
    k   = sample(1:N,size=N,replace=TRUE,prob=weight)
    x1   = rnorm(N,x[k],tau)
    lw  = dnorm(data[t],x1,sigma,log=TRUE)-dnorm(data[t],x[k],sigma,log=TRUE)
    w   = exp(lw)
    w   = w/sum(w)
    ESS  = 1/sum(w^2)

    if(ESS<(N/r)){
      index<-sample(N,size=N,replace=T,prob=w)
      x1<-x1[index]
      w<-rep(1/N,N)
    }else{}

    x <- x1
    xs = rbind(xs,x)
    ws = rbind(ws,w)
    ess =rbind(ess,ESS)

  }
  return(list(xs=xs,ws=ws,ess=ess))
}
```

```{r echo=FALSE}
apf<-APFfun(y,N,m0,C0,tau,sigma)
ESS.APF<-ESSplot(df,apf)
```

```{r echo=FALSE, fig.align='center', fig.cap="a) APF Filtered States with credible interval (in red). b) Effective sample size (in black) with threshold (in yellow).", fig.height=4, fig.pos='ht', fig.width=8, message=FALSE, warning=FALSE, results=F}
FILT.APF<-Filtplot(df,apf,"")
ggarrange(FILT.APF,ESS.APF,labels=c("(a)","(b)"))
```

Let's compare the Auxiliary Particle Filter (APF) and the Bootstrap Particle Filter (BPF).

pf2<-PFfun(y,N2,m0,C0,tau,sigma)
apf2<-APFfun(y,N2,m0,C0,tau,sigma)

ggplot(dataframe,aes(x=timeframe))+
  geom_line(aes(y=mean1, col="APF", linetype="APF"))+
  geom_line(aes(y=x, col="True States", linetype="True States"))+
  geom_line(aes(y=mean2, col="BPF", linetype="BPF"))+
  #geom_ribbon(aes(ymin = mean-1.96*sd, ymax = mean+1.96*sd),
   #           fill="red",alpha=0.16) +
    scale_color_manual("",
                      values=c("APF" = "orange",
                               "True States" = "black",
                               "BPF"= "blue"))+
  scale_linetype_manual("",
                      values=c("APF" = 1,
                               "True States" = 3,
                               "BPF"=1))+
  labs(x="Time",
       y="")+
  ggtitle(title)+
  theme_bw()+
    theme(legend.direction = "horizontal", legend.position = "bottom", legend.key = element_blank(), 
        legend.background = element_rect(fill = "white", colour = "gray30")) +
  theme(plot.title = element_text(hjust = 0.5))
}

p2<-Comparetwofunplot(df,pf2,apf2,"Filtered States")

ggplot(dataframe,aes(x=timeframe))+
geom_line(aes(y=sd2, col="APF"))+
geom_line(aes(y=sd1,col="BPF"))+
  scale_color_manual("",
                      values=c("APF" = "orange",
                               "BPF"= "blue"))+
  labs(x="Time",
       y="")+
  ggtitle(title)+
  theme_bw()+
    theme(legend.direction = "horizontal", legend.position = "bottom", legend.key = element_blank(), 
        legend.background = element_rect(fill = "white", colour = "gray30"))+
  theme(plot.title = element_text(hjust = 0.5))

c2<-ComparetwoVarplot(Combined,pf2,apf2,"Standard Deviation")

```{r include=FALSE}
truex<-x
estimatedx<-matrix(NA,ncol=4,nrow=3)
colnames(estimatedx)<-c("N","Threshold","BPF","APF")
estimatedx[,1]<-c(1000,1000,1000)
estimatedx[,2]<-c(0.5,0.25,0.1)

pf2<-PFfun(y,N,m0,C0,tau,sigma,r=k)
apf2<-APFfun(y,N,m0,C0,tau,sigma,r=k)

estimatedx[i,3]<-RMSE(truex,Filtervalues(pf2)$mean)
estimatedx[i,4]<-RMSE(truex,Filtervalues(apf2)$mean)
i=i+1

}
```
